---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/",
  fig.width = 6.5,
  fig.height = 4,
  render = knitr::normal_print,
  dev.args=list(type="cairo"),
  # dev = "png",
  dpi = 300,
  retina = 1
)
library(jtools)
```

# jtools 

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version-ago/jtools)](https://cran.r-project.org/package=jtools) [![GitHub tag](https://img.shields.io/github/tag/jacob-long/jtools.svg?label=Github)](https://github.com/jacob-long/jtools) [![Total Downloads](https://cranlogs.r-pkg.org/badges/grand-total/jtools)](https://cran.r-project.org/package=jtools)
 [![Build Status](https://travis-ci.org/jacob-long/jtools.svg?branch=master)](https://travis-ci.org/jacob-long/jtools) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/jacob-long/JTools?branch=master&svg=true)](https://ci.appveyor.com/project/jacob-long/JTools) [![codecov](https://codecov.io/gh/jacob-long/jtools/branch/master/graph/badge.svg)](https://codecov.io/gh/jacob-long/jtools) 
<!-- [![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active) [![MIT License](https://img.shields.io/badge/license-MIT-blue.svg?style=flat)](https://opensource.org/licenses/MIT) -->

This package consists of a series of functions created by the author (Jacob)
to automate otherwise tedious research tasks. At this juncture, the unifying
theme is the more efficient presentation of regression analyses. There are 
a number of functions for other programming and statistical purposes as 
well. Support for the `survey` package's `svyglm` objects as well as weighted
regressions is a common theme throughout.

**Notice:** As of `jtools` version 2.0.0, all functions dealing with 
interactions (e.g., `interact_plot`, `sim_slopes`, `johnson_neyman`) have been
moved to a new package, aptly named 
[`interactions`](https://interactions.jacob-long.com).

## Installation

For the most stable version, simply install from CRAN.

```r
install.packages("jtools")
```

If you want the latest features and bug fixes then you can download from
Github. To do that you will need to have `devtools` installed if you don't 
already:

```r
install.packages("devtools")
```

Then install the package from Github.

```r
devtools::install_github("jacob-long/jtools")
```

You should also check out the
[`dev`](https://github.com/jacob-long/jtools/tree/dev) branch of this 
repository for the latest and greatest changes, but also the latest and 
greatest bugs. To see what features are on the roadmap, check the issues
section of the repository, especially the "enhancement" tag.

## Usage

Here's a synopsis of the current functions in the package:

### Console regression summaries (`summ`)

`summ` is a replacement for `summary` that provides the user several options
for formatting regression summaries. It supports `glm`, `svyglm`, and `merMod`
objects as input as well. It supports calculation and reporting of
robust standard errors via the `sandwich` package.

Basic use:

```{r}
fit <- lm(mpg ~ hp + wt, data = mtcars)
summ(fit)
```

It has several conveniences, like re-fitting your model with
scaled variables (`scale = TRUE`). You have the option to leave the
outcome variable in its original scale (`transform.response = TRUE`),
which is the default for scaled models. I'm a fan of Andrew Gelman's 2
SD standardization method, so you can specify by how many standard deviations 
you would like to rescale (`n.sd = 2`). 

You can also get variance inflation factors (VIFs) and partial/semipartial
(AKA part) correlations. Partial correlations are only available for OLS
models. You may also substitute confidence intervals in place of standard 
errors and you can choose whether to show p values.

```{r}
summ(fit, scale = TRUE, vifs = TRUE, part.corr = TRUE, confint = TRUE, pvals = FALSE)
```

Cluster-robust standard errors:

```{r}
data("PetersenCL", package = "sandwich")
fit2 <- lm(y ~ x, data = PetersenCL)
summ(fit2, robust = "HC3", cluster = "firm")
```

Of course, `summ` like `summary` is best-suited for interactive use. When
it comes to sharing results with others, you want sharper output and probably
graphics. `jtools` has some options for that, too.

### LaTeX-, Word-, and RMarkdown-friendly regression summary tables (`export_summs`)

For tabular output, `export_summs` is an interface to the `huxtable`
package's `huxreg` function that preserves the niceties of `summ`, particularly
its facilities for robust standard errors and standardization. It also 
concatenates multiple models into a single table.

```{r eval = F}
fit <- lm(mpg ~ hp + wt, data = mtcars)
fit_b <- lm(mpg ~ hp + wt + disp, data = mtcars)
fit_c <- lm(mpg ~ hp + wt + disp + drat, data = mtcars)
coef_names <- c("Horsepower" = "hp", "Weight (tons)" = "wt",
                "Displacement" = "disp", "Rear axle ratio" = "drat",
                "Constant" = "(Intercept)")
export_summs(fit, fit_b, fit_c, scale = TRUE, transform.response = TRUE, coefs = coef_names)
```
```{r echo = FALSE, results = 'asis', warning = FALSE, eval = T}
# huxtable doesn't render right for the github_document, seems to thinks it's
# in the console
fit <- lm(mpg ~ hp + wt, data = mtcars)
fit_b <- lm(mpg ~ hp + wt + disp, data = mtcars)
fit_c <- lm(mpg ~ hp + wt + disp + drat, data = mtcars)
coef_names <- c("Horsepower" = "hp", "Weight (tons)" = "wt",
                "Displacement" = "disp", "Rear axle ratio" = "drat",
                "Constant" = "(Intercept)")
e <- export_summs(fit, fit_b, fit_c, scale = TRUE, transform.response = TRUE,
                  coefs = coef_names)
huxtable::print_html(e)
```

In RMarkdown documents, using `export_summs` and the chunk option 
`results = 'asis'` will give you nice-looking tables in HTML and PDF
output. Using the `to.word = TRUE` argument will create a Microsoft Word
document with the table in it.

### Plotting regression summaries (`plot_coefs` and `plot_summs`)

Another way to get a quick gist of your regression analysis is to 
plot the values of the coefficients and their corresponding uncertainties
with `plot_summs` (or the closely related `plot_coefs`).
Like with `export_summs`, you can still get your scaled
models and robust standard errors.

```{r}
coef_names <- coef_names[1:4] # Dropping intercept for plots
plot_summs(fit, fit_b, fit_c, scale = TRUE, robust = "HC3", coefs = coef_names)
```

And since you get a `ggplot` object in return, you can tweak and theme as
you wish.

Another way to visualize the uncertainty of your coefficients is via the
`plot.distributions` argument.

```{r}
plot_summs(fit_c, scale = TRUE, robust = "HC3", coefs = coef_names, plot.distributions = TRUE)
```

These show the 95% interval width of a normal distribution for each estimate.

`plot_coefs` works much the same way, but without support for `summ` 
arguments like `robust` and `scale`. This enables a wider range of 
models that have support from the `broom` package but not for `summ`. 

### Plotting model predictions (`effect_plot`)

Sometimes the best way to understand your model is to look at the predictions
it generates. Rather than look at coefficients, `effect_plot` lets you plot
predictions across values of a predictor variable alongside the observed data. 

```{r}
effect_plot(fit_c, pred = hp, interval = TRUE, plot.points = TRUE)
```

And a new feature in version `2.0.0` lets you plot *partial residuals*
instead of the raw observed data, allowing you to assess model quality after
accounting for effects of control variables.

```{r}
effect_plot(fit_c, pred = hp, interval = TRUE, partial.residuals = TRUE)
```

Categorical predictors, polynomial terms, (G)LM(M)s, weighted data, and much
more are supported. 

### Other stuff

There are several other things that might interest you.

* `gscale`: Scale and/or mean-center data, including `svydesign` objects
* `scale_mod` and `center_mod`: Re-fit models with scaled and/or mean-centered
data
* `wgttest` and `pf_sv_test`, which are combined in `weights_tests`: Test the
ignorability of sample weights in regression models
* `svycor`: Generate correlation matrices from `svydesign` objects
* `theme_apa`: A mostly APA-compliant `ggplot2` theme
* `theme_nice`: A nice `ggplot2` theme
* `add_gridlines` and `drop_gridlines`: `ggplot2` theme-changing convenience
functions
* `make_predictions`: an easy way to generate hypothetical predicted data
from your regression model for plotting or other purposes.

Details on the arguments can be accessed via the R documentation 
(`?functionname`). 
There are now vignettes documenting just about everything you can do as well.

## Contributing

I'm happy to receive bug reports, suggestions, questions, and (most of all) 
contributions to fix problems and add features. I prefer you use the Github
issues system over trying to reach out to me in other ways. Pull requests for 
contributions are encouraged. If you are considering writing up a bug fix
or new feature, please check out the [contributing guidelines](CONTRIBUTING.md).

Please note that this project is released with a 
[Contributor Code of Conduct](CONDUCT.md). By participating in this project
you agree to abide by its terms.

## License

The source code of this package is licensed under the 
[MIT License](http://opensource.org/licenses/mit-license.php).
