---
title: "Exploring interactions in linear regression models"
author: "Jacob Long"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring interactions in linear regression models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(message=F, warning=F)
library(jtools)
```

Understanding an interaction effect in a linear regression model is usually difficult
when using just the basic output tables and looking at the coefficients. The `jtools` 
package provides two functions that can help analysts probe more deeply.

## Simple slopes analysis

Simple slopes analysis gives researchers a way to express the interaction effect in
terms that are easy to understand to those who know how to interpret direct effects in
regression models. This method is designed for **continuous variable by continuous 
variable** interactions.

In simple slopes analysis, researchers are interested in the *conditional* slope of
the focal predictor; that is, what is the slope of the predictor when the moderator
is held at some particular value? The regression output we get when including the 
interaction term tells us what the slope is when the moderator is held at zero, which 
is often not a practically/theoretically meaningful value. To better understand the
nature of the interaction, simple slopes analysis allows the researcher to specify 
meaningful values at which to hold the moderator value.

While the computation behind doing so isn't exactly rocket science, it is inconvenient 
and prone to error. The `sim_slopes()` function from `jtools` accepts a regression model
(with an interaction term) as an input and automates the simple slopes procedure. The 
function will, by default, do the following:

* Mean-center all non-focal predictors (so that setting them to zero means setting 
them to their mean)
* For continuous moderators, it will choose the mean as well as the mean plus/minus 
1 standard deviation as values at which to find the slope of the focal predictor.
* For categorical/binary moderators, it will find the slope of the focal predictor 
at each level of the moderator.

In its most basic use case, `sim_slopes` needs three arguments: a linear model (with 
support for `svyglm` models), the name of the focal predictor in quotations as the
argument for `pred =`, and the name of the moderator in quotations as the argument
for `modx =`. Let's go through an example.

First, we use example data from `state.x77` that is built into R. Let's look at 
the interaction model output with `j_summ()` as a starting point.

```{r}
fiti <- lm(Income ~ Illiteracy*Murder, data = as.data.frame(state.x77))
j_summ(fiti)
```
So we see a significant main effect of `Murder` on `Income` in the presence of a 
significant interaction between `Murder` and `Illiteracy`. The positive estimate 
for `Illiteracy` does not differ significantly from zero.

Now let's do the most basic simple slopes analysis:

```{r}
sim_slopes(fiti, pred = "Illiteracy", modx = "Murder")
```

So what we see in this example is that when the value of `Murder` is high, the
slope of `Illiteracy` is negative and significantly different from zero. The value
for `Illiteracy` when `Murder` is high is in the opposite direction from its coefficient
estimate for the first version of the model fit with `lm()` but this result makes 
sense considering the interaction coefficient was negative; it means that as one 
of the variables goes up, the other goes down. Now we know the effect of `Illiteracy` 
only exists when `Murder` is high.

You may also choose the values of the moderator yourself with the `modxvals =` argument.

```{r}
sim_slopes(fiti, pred = "Illiteracy", modx = "Murder", modxvals = c(0, 5, 10))
```


### Additional options

#### Conditional intercepts

Sometimes it is informative to know the conditional intercepts in addition to the 
slopes. It might be interesting to you that individuals low on the moderator have a
positive slope and individuals high on it don't, but that doesn't mean that individuals
low on the moderator will have higher values of the dependent variable. You would only
know that if you know the conditional intercept.

You can print the conditional intercepts with the `cond.int = TRUE` argument.

```{r}
sim_slopes(fiti, pred = "Illiteracy", modx = "Murder", cond.int = TRUE)
```

This example shows you that while the slope associated with `Illiteracy` is negative
when `Murder` is high, the conditional intercept is also high when `Murder` is high. 
That tells you that increases in `Illiteracy` for high-`Murder` observations will 
tend towards being equal on `Income` to observations with lower values of `Murder`.

#### Robust standard errors

Certain models require heteroskedasticity-robust standard errors. To be consistent 
with the reporting of heteroskedasticity-robust standard errors offered by `j_summ()`,
`sim_slopes()` will do the same with the use of the `robust = TRUE` option so you 
can consistently report standard errors across models.

```{r}
sim_slopes(fiti, pred = "Illiteracy", modx = "Murder", robust = TRUE)
```
These data are a relatively rare case in which the robust standard errors are even 
more efficient than typical OLS standard errors. Note that you must have `sandwich` 
and `lmtest` packages installed to use this feature.

#### Centering variables

By default, all non-focal variables are mean-centered. You can additionally specify 
that the focal predictors are centered with `centered = "all"`. You may also request
that no variables be centered with `centered = "none"`. You may also request specific variables to 
center by providing a vector of quoted variable names---no others will be centered 
in that case.

Note that the moderator is centered around the specified values. Factor variables 
are ignored in the centering process.

## Plotting interactions

An even more versatile and sometimes more interpretable method for understanding 
interaction effects is via plotting. `jtools` provides `interact_plot` as a relatively
pain-free method to get good-looking plots of interactions using `ggplot2` on the 
backend.

```{r}
interact_plot(fiti, pred = "Illiteracy", modx = "Murder")
```
By default, with a continuous moderator you get two lines---1 standard deviation 
above and below the mean. If you specify `modxvals = "mean-plus-minus"`, the mean
of the moderator is also plotted.

```{r}
interact_plot(fiti, pred = "Illiteracy", modx = "Murder", modxvals = "mean-plus-minus")
```

However, if your moderator is **a factor**, each level will be plotted and you should
leave `modxvals = NULL`, the default.

```{r}
fitiris <- lm(Petal.Length ~ Petal.Width*Species, data = iris)
interact_plot(fitiris, pred = "Petal.Width", modx = "Species")
```

### Plotting points

If you want to see the individual data points plotted to better understand how the 
fitted lines related to the observed data, you can use the `plot.points = TRUE` argument.

```{r}
interact_plot(fiti, pred = "Illiteracy", modx = "Murder", plot.points = TRUE)
```
This isn't especially informative for a continuous moderator. It can be very 
enlightening, though, for categorical moderators.

```{r}
interact_plot(fitiris, pred = "Petal.Width", modx = "Species", plot.points = TRUE)
```

### Confidence bands

Another way to get a sense of the precision of the estimates is by plotting 
confidence bands. To get started, just set `interval = TRUE`. To decide how wide
the confidence interval should be, express the percentile as a number, e.g., 
`int.width = 0.8` corresponds to an 80% interval.

```{r}
interact_plot(fiti, pred = "Illiteracy", modx = "Murder", interval = TRUE, int.width = 0.8)
```

### Other options

There are a number of other options not mentioned, many relating to the appearance.

For instance, you can manually specify the axis labels, add a main title, and so on.

```{r}
interact_plot(fiti, pred = "Illiteracy", modx = "Murder", x.label = "Custom X Label",
              y.label = "Custom Y Label", main.title = "Sample Plot", 
              legend.main = "Custom Legend Title")
```

Because the function uses `ggplot2`, it can be modified and extended like any other
`ggplot2` object. For example, using the `theme_apa()` function from `jtools`:

```{r}
interact_plot(fitiris, pred = "Petal.Width", modx = "Species") + theme_apa()
```

