---
title: "j_summ --- An alternative to summary for regression models"
author: "Jacob Long"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{j_summ() — An alternative to summary() for regression models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(message=F, warning=F)
library(jtools)
```

## 

When sharing analyses with colleagues unfamiliar with R, I found that the output
generally was not clear to them. Things were even worse if I wanted to give
them information  that is not included in the output like VIFs, robust standard
errors, or standardized coefficients since the functions for estimating these
don't append them to a typical regression table. After creating output tables
"by hand" on multiple occasions, I thought it best to pack things into a
reusable function.

With no user-specified arguments except a fitted model, the output of `j_summ()` looks
like this:

```{r}
# Fit model
fit <- lm(Income ~ Frost + Illiteracy + Murder, data = as.data.frame(state.x77))
j_summ(fit)
```

Like any output, this one is somewhat opinionated—some information is shown that
perhaps not everyone would be interested in, some may be missing. That,
of course, was the motivation behind the creation of the function; the author
was no fan of `summary()` and its lack of configurability. 

## Adding and removing written output

Much of the output with `j_summ()` can be removed while there are several other pieces
of information under the hood that users can ask for. 

To remove the written output at the beginning, set `model.info = FALSE` and/or 
`model.fit = FALSE`.

```{r}
j_summ(fit, model.info = FALSE, model.fit = FALSE)
```

Another, related bit of information available before the coefficient table relates
to model assumptions (for OLS linear regression). When `model.check = TRUE`,
`j_summ()` will report (with the help fo the `car` package) two quantities
related to linear regression assumptions:

* The result of the Breusch-Pagan test for heteroskedasticity. A low *p* value
indicates the presence of heteroskedasticity.
* A count of high-leverage observations using Cook's Distance. The threshold for 
what is considered high is equal to $4/N$, where $N$ is the sample size.

In both cases, you shouldn't treat the results as proof of meaningful problems (or
a lack of meaningful problems), but instead as a heuristic for more probing with 
graphical analyses.

```{r}
j_summ(fit, model.check = TRUE)
```

## Report robust standard errors

One of the problems that originally motivated the creation of this function was the
desire to efficiently report robust standard errors—while it is easy enough for an
experienced R user to calculate robust standard errors, there are not many simple ways
to include the results in a regression table as is common with the likes of
Stata, SPSS, etc.

Robust standard errors require the user to have both `lmtest` and `sandwich` packages
installed. They do not need to be loaded.

There are multiple types of robust standard errors that you may use, ranging from 
"HC0" to "HC5". Per the recommendation of the authors of the `sandwich` package, the
default is "HC3". Stata's default is "HC1", so you may want to use that if your goal
is to replicate Stata analyses.

```{r}
j_summ(fit, robust = TRUE, robust.type = "HC3")
```

Robust standard errors can also be calculated for generalized linear models
(i.e., `glm` objects) though some debate whether they should be used for models
fit iteratively with non-normal errors. In the case of `svyglm`, the standard
errors that package calculates are already robust to heteroskedasticity,
so a `robust = TRUE` parameter will be ignored with a warning.

You may also specify with `cluster` argument the name of a variable in the input
data or a vector of clusters to get cluster-robust standard errors.


## Standardized beta coefficients

Some prefer to use standardized coefficients in order to avoid dismissing an effect
as "small" when it is just the units of measure that are small. Standardized betas
are used instead when `standardize = TRUE`. To be clear, since the meaning of
"standardized beta" can vary depending on who you talk to, this option mean-centers
the variables too.

```{r}
j_summ(fit, standardize = TRUE)
```

You can also choose a different number of standard deviations to divide by for
standardization. Andrew Gelman has been a proponent of dividing by 2 standard
deviations; if you want to do things that way, give the argument `n.sd = 2`. 

```{r}
j_summ(fit, standardize = TRUE, n.sd = 2)
```

Note that this is achieved by refitting the model. If the model took a long time
to fit initially, expect a similarly long time to refit it.

### Mean-centered variables

In the same vein as the standardization feature, you can keep the original scale
while still mean-centering the predictors with the `center = TRUE` argument.

```{r}
j_summ(fit, center = TRUE)
```

## Confidence intervals

In many cases, you'll learn more by looking at confidence intervals than 
p-values. You can request them from `j_summ`.

```{r}
j_summ(fit, confint = TRUE, digits = 2)
```

You can adjust the width of the confidence intervals, which are by default
95% CIs.

```{r}
j_summ(fit, confint = TRUE, ci.width = .5, digits = 2)
```

You might also want to drop the p-values altogether.

```{r}
j_summ(fit, confint = TRUE, pvals = FALSE, digits = 2)
```

Note that you can omit p-values without using confidence intervals.

## Generalized and Mixed models

`j_summ` has been expanding its range of supported model types. `glm` was a
natural extension and will cover most use cases.

```{r}
fitg <- glm(vs ~ drat + mpg, data = mtcars, family = binomial)

j_summ(fitg)
```

For exponential family models, especially logit and Poisson, you may be
interested in getting odds ratios rather than the linear beta estimates. 
`j_summ` can handle that!

```{r}
j_summ(fitg, odds.ratio = TRUE)
```

Standard errors are omitted for odds ratio estimates since the confidence
intervals are not symmetrical.

You can also get summaries of `merMod` objects, the mixed models from the 
`lme4` package.

```{r message = FALSE, warning = FALSE}
library(lme4)
fm1 <- lmer(Reaction ~ Days + (Days | Subject), sleepstudy)

j_summ(fm1)
```

Note the omission of p-values by default. There's no clear-cut way to derive
p-values with mixed models, since relying on the t statistics can be misleading
unless the number of subjects and groups is large. Confidence interval are 
better, but not perfect. See lme4's documentation for more info.

## Other options

### Choose how many digits past the decimal to round to

With the `digits =` argument, you can decide how precise you want the outputted 
numbers to be. It is often inappropriate or distracting to report quantities with
many digits past the decimal due to the inability to measure them so precisely or 
interpret them in applied settings. In other cases, it may be necessary to use more
digits due to the way measures are calculated.

The default argument is `digits = 3`.

```{r}
j_summ(fit, model.info = FALSE, digits = 5)
```

```{r}
j_summ(fit, model.info = FALSE, digits = 1)
```

You can pre-set the number of digits you want printed for all `jtools` functions
with the `jtools-digits` option.

```{r}
options("jtools-digits" = 2)
j_summ(fit, model.info = FALSE)
```

```{r echo = F}
options("jtools-digits" = NA)
```

Note that the return object has non-rounded values if you wish to use them later.

```{r}
j <- j_summ(fit, digits = 3)

j$coeftable
```

### Calculate and report variance inflation factors (VIF)

When multicollinearity is a concern, it can be useful to have VIFs reported alongside
each variable. This can be particularly helpful for model comparison and checking for
the impact of newly-added variables. To get VIFs reported in the output table, just
set `vifs = TRUE`.

Note that the `car` package is needed to calculate VIFs.

```{r}
j_summ(fit, vifs = TRUE)
```

There are many standards researchers apply for deciding whether a VIF is too large.
In some domains, a VIF over 2 is worthy of suspicion. Others set the bar higher, at 
5 or 10. Ultimately, the main thing to consider is that small effects are more likely
to be "drowned out" by higher VIFs.



